---
title: "Portfolio 3"
author: "Cynthia Jiao"
date: "03/07/2025"
output: html_document
---

I have my full thesis data ready, so I am going to use the analysis/graphs as my portfolio pieces.

```{r load-packages, message=FALSE}
library(tidyverse)
library(dplyr)
library(readxl)
library(tidyr)
library(lme4)
library(car)
library(emmeans)
library(effectsize)
library(irr)
library(tidycensus)
library(lme4)
library(gtsummary)
library(ggpubr)
library(tibble)
library(sf)
library(ggplot2)

```



1. Cleaning & Merging the data sets
```{r, echo=FALSE}

qualtrics <-read.csv("Cynthia Thesis_March 5, 2025_11.29.csv")

## cleaning the data by PID

qualtrics <- qualtrics %>% 
  select(-c(StartDate, EndDate, Status, IPAddress, Progress, Duration..in.seconds., RecordedDate, ResponseId, RecipientFirstName, RecipientLastName, RecipientEmail, ExternalReference, LocationLatitude, LocationLongitude, DistributionChannel, UserLanguage)) %>%
  mutate(PID = as.numeric(PID)) %>%
  arrange(PID) %>%
  filter(!is.na(PID))

## loading the data sheet from Google sheets

sheets <-read_excel("Intersectional Stereotypes Data.xlsx")

sheets <- sheets %>%
  filter(!is.na(PID)) %>%
  mutate(PID = as.numeric(PID))

## merging two data set by PID

thesis <- full_join(qualtrics, sheets, by = "PID")

thesis <- thesis %>%
  filter(!is.na(smart_screener1))

## exclude the excluded participants and participants that only have half data because of failure to complete the other training...

thesis <- thesis %>%
  filter(exclude %in% c("no", "no; half data"))



```

2. Transforming the variables

In the linear mixed-effect model, choice score of each of the 6 groups will be the outcome variable, and target race, gender, stereotype dimensions, and within-subject variance will be regressed on choice scores. To fit data into the model, I will convert the data to long format. 
```{r, echo = F}

## creating a composite score for each social group being selected

## manually specify column names for smart and nice selections

smart_cols <- c("S1_1", "S1_2", "S1_3", "S1_4", "S1_5", "S1_6", "S1_7", 
                "S1_8", "S1_9", "S1_10", "S1_11", "S1_12", "S1_13", "S1_14", "S1_15", "S2_1", "S2_2", "S2_3", "S2_4", "S2_5", "S2_6", "S2_7", 
                "S2_8", "S2_9", "S2_10", "S2_11", "S2_12", "S2_13", "S2_14", "S2_15")

nice_cols <- c("N1_1", "N1_2", "N1_3", "N1_4", "N1_5", "N1_6", "N1_7", 
               "N1_8", "N1_9", "N1_10", "N1_11", "N1_12", "N1_13", "N1_14", "N1_15", "N2_1", "N2_2", "N2_3", "N2_4", "N2_5", "N2_6", "N2_7", 
               "N2_8", "N2_9", "N2_10", "N2_11", "N2_12", "N2_13", "N2_14", "N2_15")

## create a function to count the numbers of each social group being selected

count_selections <- function(data, cols) {
  data %>%
    select(PID, all_of(cols)) %>%
    mutate(CompletedVersion = ifelse(rowSums(!is.na(select(., starts_with("S1")))) > 0, "S1", "S2")) %>%
    pivot_longer(cols = -c(PID, CompletedVersion), names_to = "Question", values_to = "Selection") %>%
    filter(!is.na(Selection) & Selection != "") %>%  # Remove unassigned version (empty values)
    group_by(PID, Selection) %>%
    summarise(Count = n(), .groups = "drop") %>%
    pivot_wider(names_from = Selection, values_from = Count, values_fill = list(Count = 0))
}

## process smart and nice data separately

smart_counts <- count_selections(thesis, smart_cols)
nice_counts <- count_selections(thesis, nice_cols)


```


3. Comparing descriptive statistics

Before fitting the variables into model, I want to first look at the mean score of each group by stereotype dimension so that I can know if there might be salient differences between groups, and more importantly, which race group to choose as the reference group in the model later.

```{r, echo = F}

# Add stereotype dimension label
smart_long <- smart_counts %>%
  mutate(stereotype_dimension = "Smart") %>%
  pivot_longer(cols = -c(PID, stereotype_dimension), names_to = "targetGroup", values_to = "choice_score")

nice_long <- nice_counts %>%
  mutate(stereotype_dimension = "Nice") %>%
  pivot_longer(cols = -c(PID, stereotype_dimension), names_to = "targetGroup", values_to = "choice_score")

# Combine into one dataset
thesis_long <- bind_rows(smart_long, nice_long)

# Extract targetGender and targetRace from "targetGroup" column

thesis_long <- thesis_long %>%
  mutate(targetGender = case_when(grepl(" man", targetGroup) ~ "Man", ##### here was the bug!!!!! fixed by adding a space in front of "man", because grepl might recognize "woman" as "man" because "woman" has "man" string too!!
                                  grepl("woman", targetGroup) ~ "Woman")) %>% 
  mutate(targetRace = case_when(
           grepl("Black", targetGroup) ~ "Black",
           grepl("White", targetGroup) ~ "White",
           grepl("Asian", targetGroup) ~ "Asian"
         ))

# Check structure
print(head(thesis_long))

## set each level as factors

thesis_long <- thesis_long %>%
  mutate(
    targetGender = factor(targetGender, levels = c("Man", "Woman")),  # Keep both levels
    targetRace = factor(targetRace, levels = c("Black", "White", "Asian")),  # Ensure 3 levels
    stereotype_dimension = factor(stereotype_dimension, levels = c("Smart", "Nice"))  # Ensure 2 levels
  )

# Descriptive statistics

thesis_long %>%
  group_by(stereotype_dimension, targetGroup) %>%
  summarise(meanChoice = mean(choice_score), sdChoice = sd(choice_score))

thesis_long %>%
  group_by(targetGender) %>%
  summarise(meanChoice = mean(choice_score), sdChoice = sd(choice_score))

thesis_long %>%
  group_by(targetRace) %>%
  summarise(meanChoice = mean(choice_score), sdChoice = sd(choice_score))

thesis_long %>%
  group_by(stereotype_dimension) %>%
  summarise(meanChoice = mean(choice_score), sdChoice = sd(choice_score))



```

From the descriptive statistics, it seems that women across race groups are chosen more than men targets. Also, Asians seemed to have the highest smart/nice scores (2.94) out of all 3 race groups, and White targets seemed to have the mean score of each dimensions (2.38) that are in between Asian and Black targets (2.18). Therefore, in the formal analysis, I will use White as the reference group. 

4. Fitting the model

I ran the model, and there are some interesting results! However, I am not quite sure if this is correct, because r is telling me that the interpret for between-subject variance is not converging for the model... I tried to but am still figuring out why it gives this message. I will continue the result interpretation as well as graphing the effects when I figure this out, hopefully in portfolio piece 4!
```{r, echo=FALSE}

# Start with white people as the race reference group given that we know that children from most racial backgrounds show a bias favoring white people
thesis_long$whiteBlack <- dplyr::recode(thesis_long$targetRace, 
                       "White" = 0,
                       "Asian" = 0,
                       "Black" = 1)

thesis_long$whiteAsian <- dplyr::recode(thesis_long$targetRace, 
                       "Asian" = 1,
                       "Black" = 0,
                       "White" = 0)


# Analysis code from preregistration

mod <- lmer(choice_score ~ targetGender * (whiteBlack + whiteAsian) * stereotype_dimension + (1 | PID), data = thesis_long)
summary(mod) ## Here is where you will get the betas (Estimate column), but ignore the p value and other stats in this code because it's not using the type of estimation parameters that you'll use

Anova(mod, type=3, test = "F") #Here you are going to get the F-statistics & p values

```

