---
title: "Portfolio 8"
author: "Cynthia Jiao"
date: "2025-4-14"
output: html_document
---

```{r setup, include = F}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(robotstxt)
library(tidyr)
library(rvest)
library(scales)
library(ggplot2)
```


# Optional Lab

Hi Mason, please see my r script in p08.R. There were just a few minor issues, but otherwise pretty smooth!

Here are my feedbacks:

1. The instructions on how to scrape the Country of Origin/Parent Company are not super clear, because the codes from lab 08 don’t entirely work here, I tried multiple times and the country_parent returns as empty (it doesn’t happen to other variables though). My guess is that the SelectorGadget did not get me the correct path, despite me selecting all the (seemingly) correct elements, so I asked GPT and eventually solved it by scraping all the rows and then only selecting rows from the 2nd column. Although I am still not sure why the old codes didn’t work, I feel like the new codes are definitely an interesting alternative to scrape data. 

2. I didn’t use str_squish as suggested, because I don’t know how str_squish could clean up the data (I know in lab 8 str_squish was used to delete the unnecessary spaces, but there was no extra spaces in this data?). I guess instead of just saying "clean it up with str_squish()", the assignment could ask:

“Some scraped text may contain irregular spacing. Use str_squish() to normalize it before analysis. Why might this be helpful?”
Or:
“After extracting the Country of Origin/Parent Company, use str_squish() and compare it with the raw version. Does anything change? When might str_squish() be necessary in web scraping?” These alternative instructions are provided by GPT!


```{r}

pac_all <-read_csv("PAC-all.csv")

```


## Data Cleaning

pac_all has 629 observations with 6 variables. Each obs represents one foreign connected PAC. 
```{r}

## separate counrty and parent company in Country of Origin/Parent Company

pac_all <- pac_all %>%
  separate(
    `Country of Origin/Parent Company`, 
    into = c("country", "parent_company"), 
    sep = "/", 
    fill = "right", 
    extra = "merge"
  ) %>%
  mutate(
    country = str_squish(country),
    parent_company = str_squish(parent_company)
  )

## convert $ amount to numeric, getting rid of comma

pac_all <- pac_all %>%
  mutate(
    Total = str_remove_all(Total, "[$,]") %>% as.numeric(),
    Dems = str_remove_all(Dems, "[$,]") %>% as.numeric(),
    Repubs = str_remove_all(Repubs, "[$,]") %>% as.numeric()
  )

```


## Data Visualization and Interpretation

#### Create a line plot of total contributions from all foreign-connected PACs in the Canada and Mexico over the years.

From the graph, it seems like that total PAC contributions from Mexico is pretty stable over years, with a small peak around year of 2016-2017. Canada, however, have skyrocketed their PAC contributions in the recent decade. Also the differences in total contribution between Canada and Mexico seem to be drastic after year of 2010.
```{r}

pac_graph <-pac_all %>%
  filter(country %in% c("Canada", "Mexico"))

## calculate total contribution of each country by year
can_mex_summary <- pac_graph %>%
  group_by(year, country) %>%
  summarize(total_contributions = sum(Total, na.rm = TRUE), .groups = "drop")

can_mex_summary

## graph the total contribution by year (x axis)

ggplot(can_mex_summary, aes(x = year, y = total_contributions, color = country)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  labs(
    title = "Total PAC Contributions from Canada and Mexico by Year",
    x = "Year",
    y = "Total Contributions (USD)",
    color = "Country"
  ) +
  scale_y_continuous(labels = scales::dollar_format()) +
  theme_minimal()



```

#### Recreate the graph

The graph shows that UK-connected PACs have consistently contributed more to Republican candidates than Democrats over the years. Although Democratic contributions had a sharp rise around 2010, they have generally remained lower compared to Republican contributions, which peaked around 2020 but declined noticeably by 2024.
```{r}

pac_uk <-pac_all %>%
  filter(country == "UK")
## calculate total contribution of the year by party

pac_uk_summary <- pac_uk %>%
  group_by(year) %>%
  summarize(
    Democrat = sum(Dems, na.rm = TRUE),
    Republican = sum(Repubs, na.rm = TRUE),
    .groups = "drop"
  )

## the graph needs a long format for data, so convert it

pac_uk_long <- pac_uk_summary %>%
  pivot_longer(
    cols = c(Democrat, Republican),
    names_to = "party",
    values_to = "amount"
  )

## recreate the graph

ggplot(pac_uk_long, aes(x = year, y = amount, color = party)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  labs(
    title = "Contribution to US politics from UK-Connected PACs",
    subtitle = "By party, over time",
    x = "Year",
    y = "Amount",
    color = "Party"
  ) +
  scale_y_continuous(labels = dollar_format(scale = 1e-6, suffix = "M")) +
  scale_x_continuous(limits = c(2000, 2024)) +
  scale_color_manual(values = c("Democrat" = "blue", "Republican" = "red")) +
  theme_minimal()



```

