---
title: "Portfolio 6"
author: "Cynthia Jiao"
date: "4/3/2025"
output: html_document
---
I will be doing the last exploratory analyses of my thesis data for this portfolio. This is a zip code analysis of the effect of neighborhood diversity (proportion of minority population) on children's stereotyping choices (spolier alert it's not significant), plus a map showing the participants' zip code distribution.

```{r setup, include = F}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(readxl)
library(tidyr)
library(lme4)
library(car)
library(emmeans)
library(effectsize)
library(irr)
library(tidycensus)
library(lme4)
library(gtsummary)
library(ggpubr)
library(tibble)
library(sf)
library(viridis)
library(ggplot2)
library(plotly)
```

# Cleaning the dataset

```{r, echo = F}

qualtrics <-read.csv("Cynthia Thesis_March 5, 2025_11.29.csv")

## cleaning the data by PID

qualtrics <- qualtrics %>% 
  select(-c(StartDate, EndDate, Status, IPAddress, Progress, Duration..in.seconds., RecordedDate, ResponseId, RecipientFirstName, RecipientLastName, RecipientEmail, ExternalReference, LocationLatitude, LocationLongitude, DistributionChannel, UserLanguage)) %>%
  mutate(PID = as.numeric(PID)) %>%
  arrange(PID) %>%
  filter(!is.na(PID))

## loading the data sheet from Google sheets

sheets <-read_excel("Intersectional Stereotypes Data.xlsx")

sheets <- sheets %>%
  filter(!is.na(PID)) %>%
  mutate(PID = as.numeric(PID))

## merging two data set by PID

thesis <- full_join(qualtrics, sheets, by = "PID")

## excluding the duplicated PID with empty Qualtrics responses

##TODO: remove remove this line for full data

## Identify PIDs that appear more than once

# duplicate_pids <- thesis %>%
#   count(PID) %>%
#   filter(n > 1) %>%
#   pull(PID) ##TODO: remove remove this line for full data

thesis <- thesis %>%
  filter(!is.na(smart_screener1))

## exclude the excluded participants

thesis <- thesis %>%
  filter(exclude %in% c("no", "no; half data"))


```

```{r, echo = F}

## creating a composite score for each social group being selected

## manually specify column names for smart and nice selections

smart_cols <- c("S1_1", "S1_2", "S1_3", "S1_4", "S1_5", "S1_6", "S1_7", 
                "S1_8", "S1_9", "S1_10", "S1_11", "S1_12", "S1_13", "S1_14", "S1_15", "S2_1", "S2_2", "S2_3", "S2_4", "S2_5", "S2_6", "S2_7", 
                "S2_8", "S2_9", "S2_10", "S2_11", "S2_12", "S2_13", "S2_14", "S2_15")

nice_cols <- c("N1_1", "N1_2", "N1_3", "N1_4", "N1_5", "N1_6", "N1_7", 
               "N1_8", "N1_9", "N1_10", "N1_11", "N1_12", "N1_13", "N1_14", "N1_15", "N2_1", "N2_2", "N2_3", "N2_4", "N2_5", "N2_6", "N2_7", 
               "N2_8", "N2_9", "N2_10", "N2_11", "N2_12", "N2_13", "N2_14", "N2_15")

## create a function to count the numbers of each social group being selected

count_selections <- function(data, cols) {
  data %>%
    select(PID, all_of(cols)) %>%
    mutate(CompletedVersion = ifelse(rowSums(!is.na(select(., starts_with("S1")))) > 0, "S1", "S2")) %>%
    pivot_longer(cols = -c(PID, CompletedVersion), names_to = "Question", values_to = "Selection") %>%
    filter(!is.na(Selection) & Selection != "") %>%  # Remove unassigned version (empty values)
    group_by(PID, Selection) %>%
    summarise(Count = n(), .groups = "drop") %>%
    pivot_wider(names_from = Selection, values_from = Count, values_fill = list(Count = 0))
}

## process smart and nice data separately

smart_counts <- count_selections(thesis, smart_cols)
nice_counts <- count_selections(thesis, nice_cols)


```

```{r, echo = F}

# Add stereotype dimension label
smart_long <- smart_counts %>%
  mutate(stereotype_dimension = "Smart") %>%
  pivot_longer(cols = -c(PID, stereotype_dimension), names_to = "targetGroup", values_to = "choice_score")

nice_long <- nice_counts %>%
  mutate(stereotype_dimension = "Nice") %>%
  pivot_longer(cols = -c(PID, stereotype_dimension), names_to = "targetGroup", values_to = "choice_score")

# Combine into one dataset
thesis_long <- bind_rows(smart_long, nice_long)


```

```{r, echo = F}

# Extract targetGender and targetRace from "targetGroup" column

thesis_long <- thesis_long %>%
  mutate(targetGender = case_when(grepl(" man", targetGroup) ~ "Man", ##### here was the bug!!!!! fixed by adding a space in front of "man", because grepl might recognize "woman" as "man" because "woman" has "man" string too!!
                                  grepl("woman", targetGroup) ~ "Woman")) %>% 
  mutate(targetRace = case_when(
           grepl("Black", targetGroup) ~ "Black",
           grepl("White", targetGroup) ~ "White",
           grepl("Asian", targetGroup) ~ "Asian"
         ))

```

```{r, echo = F}
### Center and recode variables for analysis

## center target gender predictor  

thesis_long$targetGenderC <- dplyr::recode(thesis_long$targetGender, 
                       "Man" = -.5,
                       "Woman" = .5)

## center dimensions
thesis_long$dimensionC <- dplyr::recode(thesis_long$stereotype_dimension, 
                       "Smart" = -.5,
                       "Nice" = .5)

# Two contrasts for 3-level factor:

# Let's start with white people as the race reference group given that we know that children from most racial backgrounds show a bias favoring white people
thesis_long$whiteBlack <- dplyr::recode(thesis_long$targetRace, 
                       "White" = 0,
                       "Asian" = 0,
                       "Black" = 1)

 thesis_long$whiteAsian <- dplyr::recode(thesis_long$targetRace, 
                       "Asian" = 1,
                       "Black" = 0,
                       "White" = 0)


 thesis_long$blackWhite <- dplyr::recode(thesis_long$targetRace, 
                       "White" = 1,
                       "Asian" = 0,
                       "Black" = 0)

 thesis_long$blackAsian <- dplyr::recode(thesis_long$targetRace, 
                       "Asian" = 1,
                       "Black" = 0,
                       "White" = 0)


```

# Do exposure to neighborhood diversity affect children's choice? (zip code analysis)

Exposure to neighborhood diversity does not influence children's choice ( B = -0.49, F(1. 1434) = 2.25, p = 0.13). Also no significant interaction effect involving neighborhood diversity was found (all ps > 0.1). 
```{r, echo = F}

##importing zip code from qualtrics and clean/merge with thesis data

## import Qualtrics demographic forms

demo <-read.csv("CARE Lab Demographics Questionnaire_LookIt_March 21, 2025_18.19.csv")

demo <- demo %>%
  filter(StudyID == "ISS") %>%
  mutate(PID = as.numeric(PID))

thesis_zip <-thesis %>%
  filter(Country == "US")
    
## merging two data set by PID

thesis_zip <- thesis_zip %>%
  left_join(demo, by = "PID")

## put it into long format

zip = thesis_zip %>%
  select(PID, neighborhood)

thesis_zip_long <- full_join(thesis_long, zip, by = "PID")


```

```{r, echo = F}


# Load ACS 5-year variables for 2022
acs_vars <- load_variables(2023, "acs5", cache = TRUE)
print(acs_vars)


acs <- get_acs(
  geography = "zcta",
  variables = c(
    total_population = "B01003_001",
    total_B = "B02001_003",
    total_AIAN = "B02001_004",
    total_A = "B02001_005",
    total_NHPI = "B02001_006",
    total_other = "B02001_007"
  ),
  year = 2023,
  survey = "acs5",
  geometry = FALSE
)

print(acs)

acs_wide <- acs %>%
  select(GEOID, variable, estimate) %>%
  pivot_wider(names_from = variable, values_from = estimate)

print(acs_wide)
head(acs_wide)
tail(acs_wide)


acs_wide <- acs_wide %>%
  mutate(
    minority_total = total_B + total_A + total_AIAN + total_NHPI + total_other,
    minority_proportion = minority_total / total_population
  )

print(acs_wide)

```


```{r, echo = F}

## since GEOID is not exactly the zip code (but they are super similar), here we will just download another crosswalk file to find corresponding zcta for each zip and then merge files to run the model....

crosswalk <-read.csv("zip_zcta_xref.csv")

crosswalk <- crosswalk %>%
  mutate(
    zcta = sprintf("%05d", as.integer(zcta)),
    zip_code = sprintf("%05d", as.integer(zip_code))
  )

# 1B. Format acs_wide: ensure GEOID is 5-digit string
acs_wide <- acs_wide %>%
  mutate(GEOID = sprintf("%05d", as.integer(GEOID)))

# 2. Merge ACS data with crosswalk
acs_cross <- left_join(acs_wide, crosswalk, by = c("GEOID" = "zcta"))


# 1C. Format thesis_long: ensure zip is 5-digit string
thesis_zip_long <- thesis_zip_long %>%
  mutate(neighborhood = sprintf("%05d", as.integer(neighborhood)))


thesis_zip_long <-thesis_zip_long %>%
  left_join(acs_cross, by = c("neighborhood" = "zip_code"))



## run the model

mod <- lmer(choice_score ~ (whiteBlack + whiteAsian) * minority_proportion * targetGenderC * dimensionC + (1 | PID), data = thesis_zip_long)
summary(mod)
Anova(mod, type=3, test = "F")

mod <- lmer(choice_score ~ (blackWhite + blackAsian) * minority_proportion * targetGenderC * dimensionC + (1 | PID), data = thesis_zip_long)
summary(mod)
Anova(mod, type=3, test = "F")


```

## creating a map with participant zip code distribution
```{r, echo = F}

zip_counts <- thesis_zip_long %>%
  group_by(neighborhood) %>%
  summarise(count = n(), .groups = "drop")

# need the library(sf) here

zcta_shapes <- get_acs(
  geography = "zcta",
  variables = "B01003_001",   # Total population variable; we only need it for the geometry
  year = 2020,
  survey = "acs5",
  geometry = TRUE
)

# Ensure GEOID in zcta_shapes and zip in zip_counts are both character strings.
zcta_shapes <- zcta_shapes %>%
  mutate(GEOID = as.character(GEOID))

zip_counts <- zip_counts %>%
  mutate(neighborhood = as.character(neighborhood))

# Merge participant counts with ZCTA geometries
zcta_map <- left_join(zcta_shapes, zip_counts, by = c("GEOID" = "neighborhood"))


ggplot(zcta_map) +
  geom_sf(aes(fill = count), color = NA) +
  scale_fill_viridis_c(
    option = "plasma",
    trans = "sqrt",
    na.value = "lightblue",     # areas not in dataset or NA will not be colored!
    limits = c(0, 20),
    breaks = c(0, 5, 10, 15, 20) # example breaks
  ) +
  labs(
    fill = "Participant Count",
    title = "Participant ZIP Code Distribution"
  ) +
  theme_void()



```
